The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-11 rational.tex
_
diagrams

\paragraph*{Trace.}
Every compact closed categoy admits a trace. For the additive case, we get
the following definition.  Given {{f : a+c <-> b+c }}, 
define {{trace+ f : a <-> b}} as:

@@

\noindent We have omitted some the commutativity and associativity shuffling to
communicate the main idea. We are given a value of type {{a}} which we embed
into {{a+0}} and then {{a + (c - c)}}. This can be reassociated into 
{{(a+c) - c}}. The component {{a+c}} which until now is just an appropriately 
tagged value of type {{a}} can be given to~{{f}} which transforms it to
something of type {{b+c}}. If the result is in the {{b}}-summand, it is produced
as the answer; otherwise the result is in the {{c}}-summand; {{eps+}} is used to
make it flow backwards to be fed to the {{eta+}} located at the beginning of the 
sequence. Iteration continues until a {{b}} is produced.

@@

Given that our language is reversible (Proposition~\ref{prop:logrev}), a
backward evaluator is relatively straightforward to implement: it simply
corresponds to. Indeed using the backward evaluator to calculate{{c v}} is
equivalent {{c^{dagger} v}} in the forward evaluator.

>>>>

\paragraph*{Trace.}
Every compact closed categoy admits a trace. For the additive case, we get
the following definition.  Given {{f : b1+b2 <-> b1+b3 }}, 
define {{trace+ f : b2 <-> b3}} as:

@@

\begin{center}
  \includegraphics{diagrams/thesis/trace_plus.pdf}
\end{center}

\noindent We have omitted some the commutativity and associativity shuffling to
communicate the main idea. We are given a value of type {{b2}} which we embed
into {{0+b2}} and then {{(-b1+b1)+b2)}}. This can be reassociated into 
{{-b1+(b1+b2)}}. The component {{b1+b2}} which until now is just an appropriately 
tagged value of type {{b2}} can be given to~{{f}} which transforms it to
something of type {{b1+b3}}. If the result is in the {{b3}}-summand, it is produced
as the answer; otherwise the result is in the {{b1}}-summand; {{eps+}} is used to
make it flow backwards to be fed to the {{eta+}} located at the beginning of the 
sequence. Iteration continues until a {{b3}} is produced.

@@

Given that our language is reversible (Proposition~\ref{prop:logrev}), a
backward evaluator is relatively straightforward to implement: it simply
corresponds to. Indeed using the backward evaluator to calculate {{c v}} is
equivalent {{c^{dagger} v}} in the forward evaluator.
